---
title: "Board Games"
author: ~
date: '2019-03-15'
slug: boardgames
categories: []
tags: 
  - R
  - visualizations
  - code
  - TidyTuesday
authors: []
---
```{r include=FALSE}
conflicted::conflict_prefer("filter", "dplyr")  # tells R to use filter from dplyr
knitr::opts_chunk$set(comment = ">")
library("knitr")
```

This is my first time participating in [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday) AND my first blog post. My love for board games finally convinced me to actually post something to my blog. 

In this post, I explore board game data from [Board Game Geek](https://boardgamegeek.com/) and dive deeper into game mechanics to identify game mechanics that frequently co-occur in games.  

I first loaded packages I knew I would be using as well as the data from the Tidy Tuesday GitHub.
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("magrittr")
pal <- wesanderson::wes_palette("FantasticFox1", type = "discrete") # set color palette

board_games <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-12/board_games.csv")
```

I wanted to first get a sense of board game ratings over time, contextualizing it by labeling some of my favorite games. 

```{r fig.align="center"}
favorites <- c("Small World", "7 Wonders", "Dominion", "Codenames", 
               "Betrayal at House on the Hill", "Citadels")

board_games %>% 
  ggplot(aes(x = year_published, y = average_rating)) +
  geom_point(alpha = .25, color = pal[3]) +
  geom_smooth(method = "loess", span = .5, color = pal[5]) + 
  ggrepel::geom_text_repel(data = . %>% filter(name %in% favorites), 
            aes(label = name),size = 4.3) + 
  geom_point(data = . %>% filter(name %in% favorites), 
             aes(x = year_published, y = average_rating), size = 2) +
  labs(x = "Year published", 
       y = "Average rating", 
       title = "Average board game ratings increasing over time") +
  bbplot::bbc_style() + 
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title=element_text(size=16,face="bold"))
```

Figuring out how to selectively label and color only some data points was a useful exercise, and I rarely think to perform additional data manipulation inside a `ggplot` geom. Because the labeled points were so close together, the labels overlapped, making them hard to read. Luckily, the `ggrepel` package came to the rescue in spacing them out a bit more. They're still a bit cramped, but much better than without it.  

I'm also trying out the `bbplot::bbc_style` plot theme for all the graphs in this post, and I'm liking how clean the plots look.

Two takeaways from this graph:  

1. As noted by the [FiveThirtyEight](https://fivethirtyeight.com/features/designing-the-best-board-game-on-the-planet/) article that prompted the use of this data for Tidy Tuesday, board game ratings are increasing over time. 
2. I seem to have pretty good taste in board games, at least based on their average ratings :) 

I also wanted to get a sense for how number of ratings and average ratings are related. **Do more frequently rated games have higher average ratings?**
```{r fig.align="center"}
board_games %>% 
  ggplot(aes(x = users_rated, y = average_rating)) +
  geom_point(alpha = .25, color = pal[3]) + 
  geom_smooth(method = "loess", span = .5) + 
  scale_x_log10() +
  ggrepel::geom_text_repel(data = . %>% filter(name %in% favorites), 
            aes(label = name),size = 4.3) + 
  geom_point(data = . %>% filter(name %in% favorites), 
             aes(x = users_rated, y = average_rating), size = 2) +
  labs(x = "Number of ratings (log)", 
       y = "Average rating", 
       title = "Average ratings increase with number of ratings") +
  bbplot::bbc_style() + 
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title=element_text(size=16,face="bold"))
```

It does seem that more frequently rated games also have higher ratings, and all of my favorite games also have a lot of ratings. 

# Game Characteristics
**What do the other game attributes look like?** Specifically I was interested in a game's expansions and mechanics, which were each stored in a single column, with multiple expansions and mechanics separated by commas. I used `splitstackshape::concat.split` to create list columns inside the tibble with those values, then counted the number of items in the list for each game to get a count of expansions and mechanics. Minimum number of players also had a long tail, with values above 5 infrequent, so I lumped those into one category. 
```{r fig.align="center"}
board_games <- board_games %>% 
  splitstackshape::concat.split("expansion", sep = ",", structure = "list") %>% 
  splitstackshape::concat.split("mechanic", sep = ",", structure = "list") %>% 
  mutate(expansion_n = map_dbl(expansion_list, function(x) sum(!is.na(x))), 
        mechanic_n = map_dbl(mechanic_list, function(x) sum(!is.na(x)))) %>% 
  mutate(min_players_fct = as.factor(case_when(min_players == 0 ~ NA_character_,
                                                  min_players == 1 ~ "1",
                                                  min_players == 2 ~ "2",
                                                  min_players == 3 ~ "3",
                                                  min_players == 4 ~ "4", 
                                                  min_players >= 5 ~ "5+",
                                                  TRUE ~ NA_character_)))
```

```{r  message=FALSE, warning=FALSE, fig.align="center"}
p1 <- ggplot(board_games, aes(x = min_players_fct)) + 
  geom_bar(fill = pal[1], stat = "count") +
  labs(x = "Minimum number of players", 
       y = "Number of games") +
  bbplot::bbc_style() + 
  theme(axis.text=element_text(size=10),
        axis.title=element_text(size=12,face="bold"))

p2 <- filter(board_games, playing_time <= 600) %>% 
  ggplot(aes(x = playing_time)) + 
  geom_histogram(fill = pal[1], bins = 15) +
  labs(x = "Play time (min.)", 
       y = "Number of games") +
  bbplot::bbc_style() + 
  theme(axis.text=element_text(size=10),
        axis.title=element_text(size=12,face="bold"))

p3 <- ggplot(board_games, aes(x = mechanic_n)) + 
  geom_histogram(fill = pal[1], bins = 20) +
  labs(x = "Number of mechanics", 
       y = "Number of games") +
  bbplot::bbc_style() + 
  theme(axis.text=element_text(size=10),
        axis.title=element_text(size=12,face="bold"))

p4 <- filter(board_games, expansion_n < 100) %>% 
  ggplot(aes(x = expansion_n)) + 
  geom_histogram(fill = pal[1], bins = 50) +
  labs(x = "Number of expansions", 
       y = "Number of games") +
  bbplot::bbc_style() + 
  theme(axis.text=element_text(size=10),
        axis.title=element_text(size=12,face="bold"))

p <- cowplot::plot_grid(p1, p2, p3, p4, nrow = 2)
title <- cowplot::ggdraw() + 
  cowplot::draw_label("Distributions of game characteristics", 
                      fontface='bold', size = 16)
cowplot::plot_grid(title, p, ncol = 1, rel_heights=c(0.1, 1))
```

The `cowplot` package helped me get all four plots into one figure. Most games require a minimum of two players, and the median play time is `r median(board_games$playing_time)` minutes. The plot only shows games <=10 hours, but there were `r sum(board_games$playing_time > 600)` board games with play time over 10 hours. These are likely a mix of games played over multiple sessions and erroneous data. 
Most games employ several game mechanics, and most games don't have any expansions, but there were `r sum(board_games$expansion_n > 100)` games with more than 100 expansions that I removed from the plot, including one game with `r max(board_games$expansion_n)` expansions!  

# Game Mechanics
I was most interested in further exploring the game mechanics, and since each game can have multiple mechanics, I used `tidyr::separate_rows` to get each mechanic on its own line. 

```{r fig.height= 9, fig.align="center"}
board_games_mech <- board_games %>% 
  tidyr::separate_rows(mechanic, sep = ",") %>% 
  filter(!is.na(mechanic)) %>% 
  group_by(mechanic) %>% 
  add_tally() %>% 
  ungroup() %>% 
  mutate(mechanic = fct_reorder(mechanic, n))

p5 <- board_games_mech %>% 
  ggplot(aes(x = mechanic)) + 
  geom_bar(stat = "count", fill = pal[5]) +
  coord_flip() + 
  labs(x = "Game mechanic",
      y = "Number of games") +
  bbplot::bbc_style() + 
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=12,face="bold"))

p6 <- board_games_mech %>% 
  ggplot(aes(x = mechanic, y = average_rating)) +
  geom_boxplot() +
  coord_flip() +  
    theme(axis.text.y = element_blank(),
        axis.title.y = element_blank()) +
  labs(y = "Average rating") +
  bbplot::bbc_style() + 
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=12,face="bold"))

pp <- cowplot::plot_grid(p5, p6, nrow = 1)
title <- cowplot::ggdraw() + cowplot::draw_label("Number of ratings and average rating of games by mechanic", fontface='bold', size = 16)
cowplot::plot_grid(title, pp, ncol = 1, rel_heights=c(0.1, 1))
```

```{r fig.align="center"}
board_games_mech %>% 
  group_by(mechanic) %>% 
  summarize(n = n(), average_rating = mean(average_rating)) %>% 
  ggplot(aes(x = n, y = average_rating)) + 
  geom_point(color = pal[5]) + 
    labs(x = "Number of games",
       y = "Average rating", 
       title = "Number of games and average rating of games by mechanic") +
  bbplot::bbc_style() + 
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=12,face="bold"),
        plot.title=element_text(size=16,face="bold"))
```

There are `r n_distinct(board_games_mech$mechanic)` different mechanics, with Dice Rolling and Hand Management being the most frequent. There is some variability in rating across mechanics, and the ratings of the games in a mechanic do not seem related to the number of games in that mechanic. Notably, Roll/Spin and Move stands out as having comparably lower ratings.  

I'm not familiar with what all of these mechanics actually mean, so I wanted to see the highest rated and most rated game in the most popular mechanics to get an idea of the type of games that use those mechanics.  

```{r}
board_games_mech %>% 
  group_by(mechanic) %>% 
  summarize(top_rated = first(name, order_by = desc(average_rating)), most_rated = first(name, order_by = desc(users_rated))) %>% 
  arrange(desc(mechanic)) %>% 
  head(10) %>% 
  kable()
```

# Board Game Mechanic Association Rules
Since games frequently have multiple mechanics, **are there game mechanics that commonly co-occur?** To answer this question, I used association rules to identify commonly co-occurring game mechanics. In order to use the `arules` package for association rule mining, the data has to be in "transactions" format. In order to get it in this format, I first restructured the data so that each mechanic was on its own line. From there, I transformed the data into a list, where each item in the list represents a single game and contains a factor array of that game's mechanic(s). This can then be converted into "transactions" format.

```{r message=FALSE, warning=FALSE}
library("arules")
library("arulesViz")

board_games_mech <- board_games %>% 
  select(game_id, mechanic) %>% 
  tidyr::separate_rows(mechanic, sep = ",") %>% 
  filter(!is.na(mechanic)) %>% 
  group_by(game_id) %>% 
  # filter(n() > 1) %>% # removing games with only one mechanic
  ungroup() %>% 
  mutate_all(as.factor)

mechanics_list <-  split(board_games_mech$mechanic, board_games_mech$game_id)
mechanics_transactions <- as(mechanics_list, "transactions")
summary(mechanics_transactions)
```

The board game data is now stored as a matrix. The density in a sparse matrix indicates the proportion of non-empty cells: `r round(summary(mechanics_transactions)@density,2)`. The most frequent items and itemset/transaction length distribution match the distributions above, showing that Dice Rolling is the most frequent mechanic, and 2852 games only have one mechanic. If I had not already done this exploratory analysis, I would usually use `itemFrequencyPlot` to visualize the most frequent mechanics. 

First, I'm using the `frequent itemsets` target to find combinations of game mechanics that commonly occur together.

```{r message=FALSE, warning=FALSE, results='hide'}
itemsets <- apriori(mechanics_transactions, 
                    parameter = list(supp=0.01, minlen = 2, maxlen=5, 
                                     target = "frequent itemsets"))
```

```{r message=FALSE, warning=FALSE}
summary(itemsets)
```

There are `r summary(itemsets)@items@Dim[1]` combinations of mechanics that occur in at least 1% of the data: `r summary(itemsets)@items@lengths[[1]]` two mechanic combinations, and `r summary(itemsets)@items@lengths[[2]]` three mechanic combinations.

One important metric for frequent itemsets is **support**, which indicates how frequently the itemset appears in the data set, or the proportion of games that contain the combination of mechanics. 

Below are the 10 mechanic combinations with the highest support. The most frequent combination of mechanics is Dice Rolling and Variable Player Powers, which occurred in about 5.5% of all games. This combination makes sense, because those two mechanics occurred with high frequency (they were the 1st and 4th highest, respectively). Interestingly, the combination of the two most frequent mechanics, Dice Rolling and Hand Management, only occurs 6th most frequently as a pair.

```{r}
DATAFRAME(itemsets) %>% 
  arrange(desc(support)) %>% 
  head(10) %>% 
  rename("mechanics" = items) %>% 
  kable(row.names = F, digits = 3)
```

Moving on to association rule mining, which gives output in the form of *if this then that* rules. For example, if a game employs mechanic X, it is likely to also employ mechanic Y.

```{r message=FALSE, warning=FALSE, results='hide'}
association_rules <- apriori(mechanics_transactions, 
                             parameter = list(supp=0.001, conf=0.8, maxlen=5))
```

```{r message=FALSE, warning=FALSE}
summary(association_rules)
```

There are `r summary(association_rules)@length` association rules that occur in at least .1% of the data. The support threshold is very low for association rules because there are so many games with only one mechanic that don't fit any association rules. 

Association rule mining employs an additional metric, **confidence**, which indicates how often the rule is true, or the proportion of games that have the left hand side (LHS) that also contain the right hand side (RHS). Crayon Rail System only appears in <1% of games, but 86% of the time a game is labeled as both Crayon Rail System and Route/Network Building, it is also labeled as Pick-up and Deliver, with this association rule occurring in 12 games.  

```{r}
DATAFRAME(association_rules) %>% 
  head(11) %>% 
  kable(row.names = F, digits = 3)
```

Below are the 10 association rules with the highest confidence. For the first 5, in 100% of the games that the combination of LHS mechanics occur, the RHS mechanic occurs as well. So 100% of the games labeled as Auction/Bidding, Stock Holding, and Tile Placement are also labeled as Route/Network Building, and this association occurs in 23 games. 

```{r}
DATAFRAME(association_rules) %>% 
  arrange(desc(confidence)) %>% 
  head(10) %>% 
  kable(row.names = F, digits = 3)
```

# Network graph of game mechanics
This shows constellations of mechanics that frequently occur together.  
```{r message=FALSE, warning=FALSE}
require(widyr)
require(igraph)
require(ggraph)
require(ggforce)

board_games %>% 
  select(name, mechanic) %>% 
  tidyr::separate_rows(mechanic, sep = ",") %>% 
  filter(!is.na(mechanic)) %>% 
  mutate_all(as.factor) %>% 
  pairwise_cor(mechanic, name, sort = T) %>% 
  filter(correlation > .15) %>% 
  graph_from_data_frame() %>% 
  ggraph() +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), repel = T) + 
  theme_void() + 
  labs(title = "Network of game mechanics that commonly occur together")
```

