---
title: Apple Music Year In Review
author: ~
date: '2019-01-08'
slug: apple-music
categories: []
tags: []
authors: []
draft: true
---

With 2018 coming to a close and many of my friends posting their Spotify year in review showing their most listened to songs, I was feeling jealous as an Apple Music user that I couldn't see my music year in review. So, in what could prove to be an embarassing foray into my very basic taste in music, I decided to make my own Apple Music year in review!

If you want to download your own Apple Music data, head to [Apple's Data and Privacy](https://privacy.apple.com/account) page, and select "Request a copy of your data." You can select what specific data you want, and they will put together a file for you to download, emailing you when it's ready. Mine took about 6 days for Apple to create the download. The specific file with your Apple Music listening history is "Apple Music Play Activity.csv".  

```{r, results = 'hide', message = F}
require(magrittr)
require(dplyr)
music <- read.csv(here::here("content/blog/data/Apple Music Play Activity.csv"), stringsAsFactors = F)
```

A quick visual inspection of the data showed that this file dates back to 2015 when I first subscribed to Apple Music, but since I'm just interested in this past year, I created a data set of only this year's data. In order to do that, I first had to convert the dates to a format R recognizes.

```{r}
music$Event.Start.Timestamp <- lubridate::ymd(substr(music$Event.Start.Timestamp, 1, 10))

music2018 <- dplyr::filter(music, Event.Start.Timestamp >= "2018-01-01" & Event.Start.Timestamp <= "2018-12-31")  %>% 
  dplyr::arrange(Event.Start.Timestamp)
```

I noticed some interesting things about how song listening is recorded. If a song is paused in the middle, or if other events occur, the song is recorded on multiple rows broken up by the time the event occurred. 

```{r}
head(music)
tail(music)
```


In order to avoid counting duplicates or songs that I started playing then skipped, I decided to only include songs that I listened to at least half of, which removed about half of the lines in the data set. 

Before removing entries, I totaled my play time to find that I listened to __`r round(sum(music2018$Play.Duration.Milliseconds, na.rm = T)/1000/60/60, 0)`__ hours of music last year.

```{r}
music2018 <- dplyr::filter(music2018, Play.Duration.Milliseconds/Media.Duration.In.Milliseconds > .5)
music2018 <- dplyr::filter(music2018, Play.Duration.Milliseconds == Media.Duration.In.Milliseconds)
```

It looks like I listened to at least half of __`r dplyr::n_distinct(music2018$Song.Name)`__ songs __`r NROW(music2018)`__ times last year, and I rang in the 2018 new year listening to __"`r music2018$Song.Name[1]`"__ by __`r music2018$Artist.Name[1]`__.  

```{r}
music2018_songs <- dplyr::group_by(music2018, Song.Name, Artist.Name) %>% 
  dplyr::summarize(N = n()) %>% 
  dplyr::arrange(-N)
```


Top songs
Listening by month
Listening by day of the week
Songs with the highest skip
Songs I listened to all the way through the most